---
title: "Assignment 5 - Meta-analysis of pitch in schizophrenia"
author: "Riccardo Fusaroli"
date: "3/7/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Building on the shoulders of giants: meta-analysis

## Questions to be answered

1. What is the current evidence for distinctive patterns of pitch mean and pitch sd in schizophrenia? Report how many papers report quantitative estimates, your method to analyze them, the estimated effect size of the difference (mean effect size and standard error for pitch mean, same for pitch sd) and forest plots representing it. 

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.



## Tips on the process to follow:

- Download the data on all published articles analyzing pitch in schizophrenia: https://www.dropbox.com/s/pmgw8wzc308so1p/Data.csv?dl=

```{r}
port5_data= read.csv("Data.csv", sep= ";") 
```


- Following the procedure in the slides calculate effect size and standard error of the effect size per each study. N.B. we focus on pitch mean and pitch standard deviation.
 . first try using lmer (to connect to what you know of mixed effects models)
 . then use rma() (to get some juicy additional statistics)

- Build a forest plot of the results (forest(model))


```{r}

#getting the effect size of mean for each study
library(metafor)


PitchRange_mean=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = port5_data)

PitchRange_sd=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = port5_data)

library(plyr)
PitchRange_mean=rename(PitchRange_mean, replace= c("yi"= "yi_mean", "vi" = "vi_mean"))
PitchRange_sd=rename(PitchRange_sd, replace= c("yi"= "yi_sd", "vi" = "vi_sd"))

PitchRange=merge(PitchRange_mean, PitchRange_sd)

#estimate overall effect

library(lmerTest)

model_mean <- lmer(yi_mean ~ 1 + (1 | Article), weights = 1/vi_mean, data=PitchRange, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

model_sd <- lmer(yi_sd ~ 1 + (1 | Article), weights = 1/vi_sd, data=PitchRange, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

summary(model_mean) #ß=0.1770 
summary(model_sd) #ß=-0.1251 

#meta analysis optimization
rma_mean <- rma(yi=yi_mean, vi=vi_mean, data = PitchRange, slab=Article) #tau^2=0.0607 (SE = 0.1133), I^2=44.28%
summary(rma_mean) #ß=0.1258
forest(rma_mean) 

rma_sd <- rma(yi=yi_sd, vi=vi_sd, data = PitchRange, slab=Article) 
summary(rma_sd) #ß=0.3526
forest(rma_sd) 

```

2. Do the results match your own analysis from Assignment 3? If you add your results to the meta-analysis, do the estimated effect sizes change? Report the new estimates and the new forest plots.

- Go back to Assignment 3, add your own study to the data table, and re-run meta-analysis. Do the results change?

```{r}
assignment=data.frame(
  
  PitchMeanControls= 141.44690 ,
  PitchMeanControlsSD= 49.23422 ,
  PitchSDControls= 25.81702,
  PitchSDControlsSD= 18.69103,
  
  PitchMeanSchizo= 149.49644, 
  PitchMeanSchizoSD= 55.55556,
  PitchSDSchizo= 22.88652 ,
  PitchSDSchizoSD= 33.01128,
  
  Article= "assignment 3",
  Year= 2017,
  SampleSizeSchizo= 75,
  SampleSizeContros= 74
)



# assignment3= as.data.frame(cbind(Article, Year, SampleSizeSchizo, SampleSizeContros, PitchMeanControls, PitchMeanControlsSD, PitchMeanSchizo, PitchMeanSchizoSD, PitchSDControls, PitchSDControlsSD,PitchSDSchizo,  PitchSDSchizoSD))


up_port5_data= rbind(port5_data, assignment)
```

```{r}
up_PitchRange_mean=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchMeanSchizo, m2i=PitchMeanControls, sd1i=PitchMeanSchizoSD, sd2i=PitchMeanControlsSD, data = up_port5_data)

up_PitchRange_sd=escalc('SMD', n1i=SampleSizeSchizo, n2i=SampleSizeContros, m1i=PitchSDSchizo, m2i=PitchSDControls, sd1i=PitchSDSchizoSD, sd2i=PitchSDControlsSD, data = up_port5_data)

library(plyr)
up_PitchRange_mean=rename(up_PitchRange_mean, replace= c("yi"= "yi_mean", "vi" = "vi_mean"))
up_PitchRange_sd=rename(up_PitchRange_sd, replace= c("yi"= "yi_sd", "vi" = "vi_sd"))

up_PitchRange=merge(up_PitchRange_mean, up_PitchRange_sd)

#estimate overall effect

library(lmerTest)

model_meany <- lmer(yi_mean ~ 1 + (1 | Article), weights = 1/vi_mean, data=up_PitchRange, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

model_sdy <- lmer(yi_sd ~ 1 + (1 | Article), weights = 1/vi_sd, data=up_PitchRange, control=lmerControl(check.nobs.vs.nlev="ignore", check.nobs.vs.nRE="ignore"))

summary(model_meany) #ß=0.1721
summary(model_sdy) #ß=-0.1197

#meta analysis optimization
rma_meany <- rma(yi=yi_mean, vi=vi_mean, data = up_PitchRange, slab=Article) 
summary(rma_meany) # ß=0.1224
forest(rma_meany) 

rma_sdy <- rma(yi=yi_sd, vi=vi_sd, data = up_PitchRange, slab=Article) 
summary(rma_sdy) #ß=0.2823
forest(rma_sdy) 


sum(port5_data$SampleSizeSchizo) #518
sum(port5_data$SampleSizeContros, na.rm = T) #216
```


***

3. Assess the quality of the literature: report and comment on heterogeneity of the studies (tau, I2), on publication bias (funnel plot), and on influential studies.

- Now look at the output of rma() and check tau and I2

```{r}
#tau2, I2 and CIs
rma_meany #tau^2=0.0208  I^2=26,15%  Q(df = 4) = 5.3130, p-val = 0.2567
confint(rma_meany,0.95) #CIs for tau2: (0.0000 , 0.6371)

rma_sdy #tau^2=3.4995  I^2=98,18%  Q(df = 6) = 129.4736, p-val < .0001
confint(rma_sdy,0.95) #CIs: (1.3901 , 18.1178)

#funne plots
funnel(rma_meany, main = "Random-effects model (mean)",xlab = "Standardized mean difference")
regtest(rma_meany,digits= 3) #not significant p = 0.454
ranktest(rma_meany,digits= 3) #not significant = no publication bias p = 0.483

funnel(rma_sdy, main = "Random-effects model (standard deviation)",xlab = "Standardized mean difference") 
regtest(rma_sdy,digits=3) #significant  p = 0.003
ranktest(rma_sdy) #?? p = 1.0000
```


```{r}
# Testing for influential studies 
#mean
influential= influence(rma_meany) 
print(influential) #Martinez et al 2015
plot(influential)

#sd
influential= influence(rma_sdy) 
print(influential) #Cohen et al. 2014
plot(influential)

#get them out, see what happens
#create new data
new_PitchRange = up_PitchRange

#remove Martinez for mean but not sd
new_PitchRange[12,13:14]=NA

#remove Cohen 2014 for sd (remove the row as it has no measurements for mean)
new_PitchRange = new_PitchRange[-10,]
```

Run rma() again to see how it looks like without the influential values

```{r}
#meta analysis optimization
rma_meanz <- rma(yi=yi_mean, vi=vi_mean, data = new_PitchRange, slab=Article) 
summary(rma_meanz) # ß= 0.2313
forest(rma_meanz) 

rma_sdz <- rma(yi=yi_sd, vi=vi_sd, data = new_PitchRange, slab=Article) 
summary(rma_sdz) #ß=-0.3754
forest(rma_sdz) 

#tau2, I2 and CIs
rma_meanz #tau^2=0  I^2=0%  Q(df = 3) = 0.6397, p-val = 0.8873
confint(rma_meanz,0.95) #CIs for tau2: 0.0000  0.1169

rma_sdz #tau^2=0.0267  I^2=31.63%  Q(df = 5) = 7.0540, p-val = 0.2167
confint(rma_sdz,0.95) #CIs: (0.0000  0.4135)


funnel(rma_meanz, main = "Random-effects model (mean)",xlab = "Standardized mean difference")
regtest(rma_meanz,digits= 3) #not significant p = 0.427
ranktest(rma_meanz,digits= 3) #not significant = no publication bias p = 0.333

funnel(rma_sdz, main = "Random-effects model (standard deviation)",xlab = "Standardized mean difference") 
regtest(rma_sdz,digits=3) #significant  p = 0.039
ranktest(rma_sdz) #not significant p = 0.2722
```


Code to access data from assignment3:
```{r}
scizo=subset(port3_data, port3_data$diagnosis == 1)
cont=subset(port3_data, port3_data$diagnosis == 0)
mean_of_mean_scizo=mean(scizo$mean)
sd_of_mean_scizo=sd(scizo$mean)
mean_of_sd_scizo=mean(scizo$sd)
sd_of_sd_scizo=sd(scizo$sd)

port5_scizo= c(mean_of_mean_scizo, sd_of_mean_scizo, mean_of_sd_scizo, sd_of_sd_scizo)
port5_scizo


mean_of_mean_cont=mean(cont$mean)
sd_of_mean_cont=sd(cont$mean)
mean_of_sd_cont=mean(cont$sd)
sd_of_sd_cont=sd(cont$sd)

port5_cont= c(mean_of_mean_cont, sd_of_mean_cont, mean_of_sd_cont, sd_of_sd_cont)
port5_cont

length(unique(scizo$id_diff)) 
length(unique(cont$id)) 

```